\iflanguage{ngerman}
{\chapter{Einleitung}}
{\chapter{Introduction}}

\label{sec:introduction}


The term \textit{"agent"} is commonly used these days. This applies to both \ac{AI} and fields outside of it, such as databases and automated manufacturing. Despite growing in popularity, the phrase has been employed in so many different contexts that it has lost all meaning when not connected to a certain understanding of agenthood. The terminology "agent" is most frequently used in the context of  \ac{AI} to describe an object that operates constantly and autonomously in a context that also contains other processes and agents. Perhaps this is the single characteristic that all practitioners of the word in \ac{AI} agree upon. Although the emphasis is on the word "autonomy,"  the definition of autonomy is vague, but it is generally understood to suggest that the behaviors of the agents do not need ongoing human direction or interference, and presumptions are frequently made about the environment of the agents, such as that it is mostly unpredictable.

\vspace{.5cm}

Agents are frequently assumed to be robotic agents, in which case additional problems including sensory input, motor control, and time constraints are stated. Finally, agents are frequently perceived as "high-level." Although this meaning is somewhat nebulous, many people use some variation of it to separate agents from other pieces of hardware or software. The "high level" is exhibited by symbolic representation and/or cognitive-like skills. Agents may be "informable," have symbolic planning in addition to stimulus-response rules, and even be able to speak. This meaning is not universally assumed in \ac{AI}, so clearly, the concept of agenthood in \ac{AI} is far from clear. As a result, We should be explicit about what we mean when we use the term "agent," which is as follows: An entity whose state is seen to be composed of mental elements including beliefs, capabilities, decisions, and commitments is referred to as an agent. These elements are precisely specified and roughly correspond to their equivalents from common sense. Therefore, according to this perspective, an agent only exists in the programmer's head. Any hardware or software component only qualifies as an agent if its analysis and management have been done in this way.

\vspace{.5cm}

In latest years, there has been a lot of interest in the research of computational agents that can behave rationally. The architecture of systems required to conduct high-level management and control activities in complex, dynamic situations is becoming increasingly important in the commercial world. Air traffic control, telecommunications networks, commercial operations, spacecraft, and medical services are examples of such systems. Experience with using traditional software methodologies to design such systems has revealed that they are extremely complicated and costly to build, validate, and sustain. Agent-oriented systems, which are based on a fundamentally different perspective of computational entities, provide opportunities for a qualitative shift in this stance. 

\vspace{.5cm}

One such architecture views the system as a rational agent with particular mental attitudes of \ac{BDI}, reflecting, respectively, the informational, motivating, and deliberative states of the agent. When deliberation is subject to resource constraints, these mental attitudes govern the system's behavior and are crucial for obtaining adequate or optimal performance.

\vspace{.5cm}

\ac{BDI} agents can be considered for both a formulation of a theory and a practical design standpoint. \textit{AgentSpeak(L)} is a generalization of one of the implemented \ac{BDI} systems that allows agent programs to be developed and processed in the same way as a horn-clause rationale programs. It is a language of programming based on a limited first-order language with events and actions. The programs created in AgentSpeak(L) govern the agent's behavior (i.e., its interaction with the environment). The agent's beliefs, desires, and intentions are not explicitly stated as modal formulae. Instead, as designers, one may attribute these ideas to AgentSpeak(L) agents. The agent's present belief state may be considered a model of itself, its environment, and other agents; the states that the agent seeks to bring about based on external or internal stimuli can be viewed as desires, and the adoption of programs to meet such stimuli can be viewed as intents.

\vspace{.5cm}

It is therefore easy to design an agent capable of controlling its surroundings on its own, but this becomes troublesome when the environment is shared by more than one agent. AgentSpeak(L) is extended by languages such as JASON and \ac{ASTRA}. They could be used to implement that language's operational semantics and provide a foundation for the development of \ac{MAS} with different user-customizable features.

\vspace{.5cm}

The implementation of these agents in specialized fields would be helpful when discussing the agents' ability to communicate using AgentSpeak(L). One of the most significant areas will be supply chain management, where we may assign roles to various agents such as manufacturers, wholesalers, and retailers. An adaptable supply chain allows retailer, distributor, and manufacturer agents to adjust to environmental changes. Each agent has a core objective as well as related secondary objectives. \ac{BDI} agents can be used to improve decision-making in supply chain management by allowing for the simulation and optimization of different scenarios and strategies. For example, a \ac{BDI} agent could be used to forecast demand for a product and adjust inventory levels accordingly. 

\vspace{.5cm}

Additionally, \ac{BDI} agents can also be used for negotiation and coordination with other agents for obtaining better prices or delivery schedules. It seems logical that whenever agents interact in the supply chain, there will be a transaction, and that transaction should be recorded. The transaction can be stored in any database by agents, but will it be optimal? A programmable ledger can be used to record the movement and ownership of items between each agent. So rather than keeping these interactions in the form of transactions in databases or spreadsheets, it will be perfect for storing them in smart contracts, which can be inspected for tracking or future references.	

\vspace{.5cm}

The concept of \textit{smart contracts}, or computer protocols intended to automatically facilitate, authenticate, and implement the negotiation and application of digital contracts without the need for centralized authorities, has been revived in recent years has resulted of the rapid development of cryptocurrencies and the \ac{BCT} that underpins them. Smart contracts have been incorporated into popular blockchain-based development platforms like Ethereum and Hyperledger. They have a wide range of potential application areas in the globalised era and intelligent sectors of the economy, including financial sector, management, healthcare, and Internet of Things (IoT), among many others. Smart contracts offer a tamper-proof and auditable record of all the transactions between various parties, which may be used to improve traceability and transparency in the supply chain. In lengthy and complicated supply chains where it is challenging to follow the flow of items, this might be very helpful.

\vspace{.5cm}

The fields of agents and blockchain seem to work well together. To find out if we can describe supply chain participants (retailer, wholesaler, and manufacturer) in Solidity language to work with the Agent \ac{BDI} Library and if agents can coordinate their actions with rules through smart contracts, we must first examine each component separately before combining them. Additionally, we will see whether we can synchronize our intentions and objectives via AgentSpeak with the payload of on-chain smart contract transactions.

\vspace{.5cm}

In this thesis, a \textbf{\textit{Smart contract development with JASON \ac{BDI} Agents}} is presented. We will attempt to develop an application that will display various agents, allowing agents across the supply to interact with one another. This ecosystem will also allow agents to maintain relationships, store transactions using smart contracts, and determine whether both the technologies can be used to create a fair and efficient dispute resolution mechanism in the supply chain by providing a transparent and tamper-proof record of the agreements and interactions between parties.

\section{Motivation}

The blockchain is having an increasing influence on a variety of research and application fields, from distributed computing and storage to supply chain management and healthcare accountability. In fact, there are many distinct use cases where the promise of supplying a secure and fault-tolerant ledger that mutually untrusted parties may use to monitor computations and interactions completely dispersed and decentralized without the need for a central authority is intriguing. 

\vspace{.5cm}

The agent community is no different; it began expressing interest in the opportunities presented by the blockchain to either solve long-standing problems (such as trust management in open systems, accountability of actions for liability, among others), or to take advantage of expected benefits to endow a system with new properties (such as novel infrastructures for \ac{MAS}, trustworthy coordination). \ac{BDI} agents may be used to build more resilient and sustainable supply chains by simulating and optimizing various sustainability scenarios, such as examining the environmental effect of different transportation routes or manufacturing processes. \ac{BDI} agents can also be programmed to automatically initiate smart contracts with suppliers or customers to enforce sustainability-related commitments.

\vspace{.5cm}

Smart contracts play a critical role in enabling the blockchain to function as a general-purpose distributed computing engine while retaining its core properties of security, trust, and fault-tolerance, and they actually extend their reach to cover computations other than data storage and management. Smart contracts can be utilized to improve supply chain traceability and transparency by providing a tamper-proof and auditable record of all transactions between multiple parties. This is especially beneficial in complicated and extensive supply chains where it is difficult to maintain track of products movement.

\vspace{.5cm}
The convergence of the agent and blockchain worlds appears to be promising: Agents are distributed autonomous entities whose interactions should be managed and mediated in trustworthy ways; blockchains and smart contracts, on the other hand, are trust-sensitive mechanisms for mediating and managing interactions. Combining \ac{MAS} with smart contracts can have several potential benefits for supply chain management and other industries:

\begin{itemize}
    
\item \textbf{Improved coordination} \\ By coordinating the activities of multiple agents through smart contracts, the supply chain can be more efficient and effective. Smart contracts can be used to define the rules and constraints that govern the interactions between agents in a supply chain, which can help to synchronize the activities of different agents and reduce the risk of errors or inconsistencies.

\vspace{.5cm}

\item \textbf{Decentralized decision-making} \\ By programming agents to make decisions based on their beliefs, desires, and intentions, multiple agent systems can provide a decentralized and autonomous approach to decision-making. This allows agents to make decisions based on the most current information and adapt to changing conditions in the supply chain.

\vspace{.5cm}

\item \textbf{Transparency} \\ Smart contracts and multiple agent systems can provide transparency in the supply chain. Smart contracts can be used to record all activities and interactions among agents, allowing all parties to see the flow of goods and the state of the supply chain.

\vspace{.5cm}

\item \textbf{Improved security} \\ By using smart contracts to define the rules and constraints that govern the interactions between agents, the system can be more resistant to malicious attacks and fraud. Additionally, multiple agents can work together to detect and defend against any malicious attempt.

\vspace{.5cm}

\item \textbf{Scalability} \\ Smart contracts can be used to handle a large number of transactions and interactions between agents, which can help to scale the system as the number of agents and transactions increases.

\vspace{.5cm}

\item \textbf{Interoperability} \\ By using smart contracts to facilitate communication and coordination among multiple agent systems, the system can be designed to be interoperable with other blockchain and non-blockchain platforms.

\vspace{.5cm}

\item \textbf{Flexibility} \\ By using multiple agents with \ac{BDI} approach, the system can adapt to changing conditions and goals, making the system more flexible.
\end{itemize}

It is crucial to remember that while merging \ac{MAS} with smart contracts might have numerous advantages, doing so takes knowledge of both disciplines, and the system's design must take into consideration the supply chain's complexity and dynamic nature.

\section{Our contributions}

The contribution of the paper are as follows:

\begin{itemize}

    \item We demonstrate and analyze the aspects of \ac{MAS}-\ac{BCT} created by integrating agents with smart contracts, as well as the implications for agent-oriented practice and blockchain. The primary goal was to utilize AgentSpeak(L) to program agents and store their interactions using smart contracts.
    
\vspace{.5cm}

    \item  The fundamental concept was to construct \ac{MAS} using JASON, an extension of AgentSpeak(L), and smart contracts using Solidity and integrating them together.
    
\vspace{.5cm}

    \item Our research enabled the usage of JASON with both the Python interpreter and Java. As an alternative to JASON, we also tested using \ac{ASTRA}. Because Java was the prior implementation language, \ac{ASTRA}'s type system is based on it. By using a comparable type system, translating between \ac{ASTRA} and Java is made easier and more clear. Vyper was also taken under consideration for Smart contracts while switching from Java to Python, but was later abandoned owing to some of its drawbacks.
    
\vspace{.5cm}

\end{itemize}

In order to integrate them, we conducted extensive study while doing that. When choosing a programming language to integrate agents written in agentSpeak(L) with smart Contracts written in Solidity, we run into a myriad of challenges.

\section{State of the Art}

Most methodologies to bridging \ac{MAS} and \ac{BCT} are of placing a \ac{MAS} and a blockchain side by side, allowing agents to utilize blockchain services opportunistically as needed. We contend that these techniques do not fully integrate agents and blockchain on a technical level, because the \ac{MAS} just exploits the blockchain like any other software library. In fact, the agents are viewed essentially like any other off-chain entity, such as blockchain clients, whose contact with the blockchain is restricted to making transactions and deploying/invoking smart contracts—just like any other service.

\vspace{.5cm}

Efforts striving to put elements of agent-oriented models and technologies directly into the blockchain.
For instance, adopting agent programming languages for implementing smart contracts would be an
integration effort belonging to this category, as well as empowering smart contracts’ computational
model (usually, object-oriented) with features defining agent-orientation. Another approach would comprise integration efforts attempting to introduce concepts and techniques from the \ac{BCT} domain into agents. An agent-programming platform, for example, can be offered, in which agents are written in a smart contract language and their interactions are enabled and governed by a \ac{BCT}'s transactions mechanism.


\vspace{.5cm}

The existing research focuses on both computational and interactional aspects. Many strategies focus on the computational aspect, aiming to use the blockchain to store data in a safe and traceable manner in order to hold agents who manipulate it accountable for their actions; other strategies concentrate on the interaction aspect, making use of the blockchain transaction mechanism, frequently in conjunction with smart contracts, to validate and control interactions among agents. However, we stress that not all aspects of the interaction dimension have been taken into account. Existing approaches always assume that interactions will take place between off-chain entities, whether they contain agents or not. However, this leaves open the possibility of using agent-oriented abstractions to mediate on-chain entity interactions, such as those involving smart contract reciprocal invocations or node communication protocols.

\vspace{.5cm}

In this thesis, we will primarily concentrate on agent-oriented models and technologies that are directly integrated into the blockchain.

\section{Evaluation Results}

Our evaluation was conducted using three separate criteria. (1) Smart Contract Functionality (2) Choosing \ac{AOP} Language (3) Outcomes of combining smart contracts with Agents. Experimental findings from these many angles are systematically examined.

\subsubsection{Smart Contract Functionality}
  
  We will check if it is possible to describe supply chain members (such as a retailer, wholesaler, and manufacturer) in Solidity, the programming language used for writing smart contracts on the Ethereum blockchain. We examined how well smart contracts run on several networks, including the local network, Infura utilizing Rinkeby, and Alfajores, as well as how long it takes to construct each contract. Local networks consistently had the lowest delay time across all the networks after testing the prototype for smart contracts, thus we utilized this network to test our subsequent scenarios involving agents. Since CELO is simpler to obtain than RinkenyETH, testing on Alfajores test network was likewise simpler than on Rinkeby test network. In none of our test scenarios did we use the mainnet.
  
  \vspace{.5cm}
  
  Along with that, we also stated reason and provided information on why Solidity is preferred over Vyper. Vyper, on the other hand, is built on Python. Solidity is similar to JavaScript, however Vyper lacks several functions that Solidity has, helping to make Solidity more efficient.

\subsubsection{Choosing \ac{AOP} Language}

  In order to understand how they operate and determine whether they are compatible with other languages to be used to include them into blockbhain, we developed the \texttt{.asl} files using JASON and \ac{ASTRA}. We utilized Java and Python based interpreters to create \ac{MAS} utilizing JASON AgentSpeak(L), and we also tested \ac{ASTRA} to learn more about the differences between the languages and see if it works well with Java and makes use of all Java packages like \texttt{org.web3j}. 

  \vspace{.5cm}
  
  Our evaluation revealed that, despite the fact that the code in the \texttt{.asl} files for JASON's Java and Python absed interpreters is identical, the two nevertheless use distinct instructions to communicate. The key difference between the two is that the Java based interpreter requires a \texttt{mas2j} file to execute \ac{MAS}, but the Python based interpreter just requires a python script that establishes an environment for multi agents to communicate with each other.

  \vspace{.5cm}

  Although \ac{ASTRA}-written agents are somewhat unique since they are more likely to resemble Java-style syntax. There is no need to create separate files for the interaction between agents because all the agents may be written in a single file with the \texttt{.astra} extension.

\subsubsection{Outcomes of smart contracts with Agents}

The principal goal of this thesis was to operate the agents in a \ac{MAS} while adapting a supply chain and collaborating \ac{BCT} to it in the form of smart contracts. We experimented with combining several AgentSpeak languages with web3 libraries for the assessment. Only \texttt{web3.py} and a Python-based interpreter for JASON were found to be effective for the implementation. With other web3 libraries, such as the \texttt{web3j} and \texttt{web3.js} libraries, we had problems with jar files and missing packages while using \ac{AOP}, since \ac{ASTRA} and java-based interpreter of JASON can't work with the core libraries required to run smart contracts.

 \vspace{.5cm}
 
In order to make it easier to grasp, the evaluation of the thesis is outlined in more detail later in the chapter.

\section{Outline of the thesis}
\begin{itemize}
    \item The research and associated work that has been done in the area of our study are covered in Chapter 2.
    
\vspace{.5cm}

    \item The technical information needed to comprehend the ideas offered in this thesis is provided in Chapter 3.
    
\vspace{.5cm}

    \item The system's high-level design, architecture, technical details and several sample use cases created for this work are covered  are covered in Chapter 4 of this thesis.
    
\vspace{.5cm}

    \item Chapter 5 summarizes the system's evaluation findings.
    
\vspace{.5cm}

    \item The design enhancements for upcoming and additional study subjects that result from this effort are covered in Chapter 6.
\end{itemize}


