{\chapter{Further Work}}
As part of our ongoing efforts, we may expand the functionality of our program and create a fantastic user interface, which will allow us to modify the code and make it usable. In order to utilize the web3j package, we also anticipate having a successful implementation with our other choices, maybe following a significant update in all the other languages we tried to employ that include a Java-based interpreter. Future work on integrating smart contracts with \ac{BDI} agents will be done in a number of domains, including:

\begin{itemize}
    \item \textbf{Better reasoning and decision-making abilities} \\ 
    One of the most important advantages of \ac{BDI} agents is their ability to make decentralized, independent decisions based on their beliefs, wants, and intents. However, existing smart contract implementations of \ac{BDI} agents may not be as complex in decision-making as anticipated. Future work in this area might include developing and implementing more complex reasoning and decision-making algorithms for \ac{BDI} agents to allow improved supply chain decision-making.

    \vspace{.5cm}
    
    \item \textbf{Better communication and coordination} \\ 
    \ac{BDI} agents can be programmed to interact and collaborate with other agents to achieve shared objectives. However, present communication and coordination channels may fall short of expectations. Future work in this area might include investigating and implementing more advanced communication and coordination methods to increase the efficacy of \ac{BDI} agents in the supply chain.

    \vspace{.5cm}
    
    \item \textbf{Merging with additional \ac{AI} technologies}\\ 
    \ac{BDI} agents can be combined with other \ac{AI} technologies such as machine learning and natural language processing. Future work in this area might include investigating and developing methods for combining \ac{BDI} agents with other \ac{AI} technologies to increase their performance and capabilities.

    \vspace{.5cm}
    
    \item \textbf{Decentralized trustworthiness and reputation} \\ 
    Agents of the \ac{BDI} can be designed to consider the standing and reliability of other agents while making judgments. The methods in place to measure reputation and trust may not be as sophisticated as they may be. Future study in this area may focus on developing and utilizing more sophisticated reputation and trust measurement techniques that can be utilized to increase the efficiency of \ac{BDI} agents in the supply chain.
    
    \vspace{.5cm}
    
    \item \textbf{More advanced simulation ecosystems} \\ 
    Agents' simulation surroundings must be upgraded when they are implemented in increasingly realistic and complicated contexts. Future work in this field might include developing and deploying more realistic and complicated simulation environments for testing and evaluating \ac{BDI} agents in real-world circumstances.

    \vspace{.5cm}
\end{itemize}
Additionally, \ac{BDI} agents are frequently intended to interact with the digital environment and may have limited interaction skills with the real world. Future work in this field might include investigating and implementing solutions for improved connection with the physical environment, such as combining \ac{BDI} agents with IoT devices to enable more effective and efficient supply chain management.