
\course{Distributed System Engineering}
\matriculationyear{2020}
\issuedate{17.10.2022}
\duedate{20.03.2023}
\professor{Prof. Dr. rer. nat. Uwe Aßmann}


\begin{task}[] % If needed, a custom title can be provided between the brackets
	\minisec{\objectivesname}\smallskip	
	Agent-oriented programming has been developed over the few decades in order to comprehend the relationship between dynamic environments and software applications. Belief-Desire-Intention agents can implement a plan execution library that consists of objectives and goals. Beliefs are the environmental status and these agents can update the status of the environment. For instance, weather degree is a dynamic environment variable that can be updated by agents to use with rule-based conditions such as desires and intentions. In order to realize agent-based abstraction, one can use the Jason Belief-Desire- Intention (BDI) agent with AgentSpeak. Jason Agent Framework is an interpreter of the logic- based BDI language called AgentSpeak so that agents can be easily integrated into a general-purpose language.
 
    \vspace{.5cm}
    
    A smart contract has preconditions and postconditions to support a handshake mechanism between participants in a blockchain network. When a transaction has been commenced in a smart contract, a transaction can update the state of the ledger. However, transactions or meta-transactions (only data) can be initiated by an externally owned account or client application. An agent can manage autonomous entities according to its internal planning library. Each plan can have multiple subgoals and subgoals can deploy existing smart contracts according to preconditions and postconditions in a case study. 
    
    \vspace{.5cm}
    
    A case study can be realized as follows: An adaptive supply chain can adapt to the environmental changes between retailer, distributor, and producer agents. Each agent has a main goal and sub-goals that are relevant to the particular main goal. Storage of retailers, distributors, producer agents can be registered into a programmable ledger. Storage capacity of retailer-distributor agents and production capacity of producer agents should be tracked by the internal planning library of agents. An agent can decide raw materials and goods can be sent according to parameters such as production ready (boolean), capacity level of storage, and production level with preconditions and postconditions. Each smart contract should be managed by agents with regards to this simulation.
    
    \vspace{.5cm}
    
    This work should investigate and identify the feasibility of decentralized smart contract execution with agents. To this end, a literature review should be provided by the thesis author regarding agent-oriented programming, contract-oriented programming, and smart contract programming. After the literature analysis, a proof-of-concept application that is relevant to a particular case study should be realized. Finally, evaluations need to be defined with corresponding criteria and the student should discuss results to come up with conclusions.
	
	\minisec{\focusname}\smallskip
	\begin{itemize}
		\item Realizing the literature review of BDI-based agent-oriented programming, decentralized smart contract execution with autonomous agents.
        
		\item Carrying out a demonstration of the feasibility of a case study.
        
		\item Comparing with other BDI agent frameworks with regards to drawbacks and advantages.
	\end{itemize}

        \minisec{Task Description}\smallskip
        The following items are possible tasks:
        \begin{enumerate}
		\item Analyzing the feasibility of the agent-oriented planning library with decentralized smart contracts.
       
		\item Researching the feasibility of the language abstraction in on-chain programming with language features of an agent such as belief-desire-intention.
        
		\item Define types of agents, which are required in a particular case study and implement the case study through the Jason-Agentspeak framework with Solidity language. 
        
        \item  In order to ensure software functionality, the student should implement behavioral and unit test cases.
	\end{enumerate}
    A demo and presentation will be organized.
\end{task}

