\TUDoption{abstract}{section, multiple}

% TODO: Replace the placeholders with your actual abstracts

\newcommand*{\Abstreng}{
   Over the last few decades, agent-oriented programming has been constituted in an attempt to better understand how software applications interact with dynamic circumstances. The modeling of intelligent systems that function in contexts with limited or ambiguous information is made particularly effective by \ac{BDI} (Belief-Desire-Intention) agents. The Jason framework, an open-source platform for building multi-agent systems in the AgentSpeak language, is a popular framework for constructing \ac{BDI} agents.
   
   \vspace{.5cm}
   
   Utilizing of blockchain-related technologies has skyrocketed in recent years too, and various blockchains customized to specific use cases have emerged. One of the most essential blockchains uses is smart contracts. Most present smart contract systems presume that when contracts are executed over a network of decentralized nodes, the majority's decision can be trusted. However, we have seen that people associated with a smart contract may intentionally take steps to manipulate contract execution in order to improve their own gains. To solve this issue, we suggest an agent model as the underlying mechanism for contract execution over a network of decentralized nodes and a public ledger, and we analyze the prospect of prohibiting users from influencing smart contract execution.
   
    \vspace{.5cm}
    
    In order to increase the automation and autonomy of a decentralized system, this thesis investigates the combination of smart contract technology with Jason \ac{BDI} agents. The paper provides an idea of the integration of the two technologies suggesting the use case of a supply chain, detailing the required elements and how they interact. Overall, the study's findings show how \ac{BDI} agents and smart contracts potentially work together to improve supply chain efficiency and transparency.}


\begin{abstract}[pagestyle=empty.tudheadings]
	\iflanguage{ngerman}
		{\Abstrger}
	
	\iflanguage{ngerman}
	{\nextabstract[english]}
	
	\iflanguage{ngerman}{
		\Abstrger}{
		\Abstreng
	}

\end{abstract}
